\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-02-20 doc/paper.tex
adding run-time comparison section

%macros
\newcommand{\sarah}{SARAH\xspace}
\newcommand{\fs}{FlexibleSUSY\xspace}
\newcommand{\ESSM}{E$_6$SSM\xspace}
\newcommand{\code}[1]{\lstinline|#1|}  % inline source code
\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}
\newcommand{\DRbar}{\textoverline{DR}}  % inline source code
\def\at{\alpha_t}
\def\ab{\alpha_b}
\def\as{\alpha_s}
\def\atau{\alpha_{\tau}}
\def\oat{O(\at)}
\def\oab{O(\ab)}
\def\oatau{O(\atau)}
\def\oatab{O(\at\ab)}
\def\oatas{O(\at\as)}
\def\oabas{O(\ab\as)}
\def\oatababq{O(\at\ab + \ab^2)}
\def\oatqatababq{O(\at^2 + \at\ab + \ab^2)}
\def\oatasatq{O(\at\as + \at^2)}
\def\oatasabas{O(\at\as +\ab\as)}
\def\oatasabasatq{O(\at\as + \at^2 +\ab\as)}
\def\oatq{O(\at^2)}
\def\oabq{O(\ab^2)}
\def\oatauq{O(\atau^2)}
\def\oabatau{O(\ab \atau)}
\def\oas{O(\as)}
\def\oatauqatab{O(\atau^2 +\ab \atau )}

@@

\subsection{Lattice method}
If ready.
\subsection{multiple effective models at different scales}
\section{Implementing new models}
\section{Adapting the code at the C++ level}
\section{Using individual modules in your own code}
\appendix
\section{Examples}
\subsection{NMSSM}
Detailed steps to create NMSSM model file.
\begin{enumerate}
\item Create new directory \code{mkdir myNMSSM}
\item Move into this new directory \code{cd myNMSSM}
\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myNMSSM.m}
\item Edit myNMSSM.m as follows
\begin{enumerate}
\item  Change Model name, author and date
\item Add singlet superfield to SuperFields
\begin{lstlisting}
SuperFields[[8]] = {s, 1, sR,    0, 1,  1, RpM};
\end{lstlisting}
\item Modify superpotential
\begin{lstlisting}
SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Kappa]/3 s.s.s ;
\end{lstlisting}
\item Add new VEV
\begin{lstlisting} 
  DEFINITION[EWSB][VEVs]= 
  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
};
\end{lstlisting}
\item  Add the singlino to the list of gauge eigenstate dirac fermions
\begin{lstlisting}
  DEFINITION[GaugeES][DiracSpinors]={
 .............
  FS -> {FsR,conj[FsR]}   
};
\end{lstlisting}
\item Edit mass eigenstates to give how the gauge eigenstates mix into mass eigenstates.  Here we add to the cp even higgs a real singlet, to the CP odd the imaginary part of the singlet and finally to the neutralinos we add the singlino.

>>>>

%macros
\newcommand{\sarah}{SARAH\xspace}
\newcommand{\fs}{FlexibleSUSY\xspace}
\newcommand{\ESSM}{E$_6$SSM\xspace}
\newcommand{\code}[1]{\lstinline|#1|}  % inline source code
\newcommand{\textoverline}[1]{$\overline{\mbox{#1}}$}
\newcommand{\DRbar}{\textoverline{DR}}  % inline source code
\newcommand{\unit}[1]{\,\text{#1}}      % units
\DeclareMathOperator{\sign}{sign}
\def\at{\alpha_t}
\def\ab{\alpha_b}
\def\as{\alpha_s}
\def\atau{\alpha_{\tau}}
\def\oat{O(\at)}
\def\oab{O(\ab)}
\def\oatau{O(\atau)}
\def\oatab{O(\at\ab)}
\def\oatas{O(\at\as)}
\def\oabas{O(\ab\as)}
\def\oatababq{O(\at\ab + \ab^2)}
\def\oatqatababq{O(\at^2 + \at\ab + \ab^2)}
\def\oatasatq{O(\at\as + \at^2)}
\def\oatasabas{O(\at\as +\ab\as)}
\def\oatasabasatq{O(\at\as + \at^2 +\ab\as)}
\def\oatq{O(\at^2)}
\def\oabq{O(\ab^2)}
\def\oatauq{O(\atau^2)}
\def\oabatau{O(\ab \atau)}
\def\oas{O(\as)}
\def\oatauqatab{O(\atau^2 +\ab \atau )}

@@

\subsection{Lattice method}
If ready.
\subsection{multiple effective models at different scales}
\section{Implementing new models}
\section{Adapting the code at the C++ level}
\section{Using individual modules in your own code}

\section{Run-time comparison with other spectrum generators}

We compare the run-time of the four different spectrum generators
FlexibleSUSY (version 0.5.3), Softsusy (version 3.4.0), SPheno
(3.2.4), SPhenoMSSM (generated with SARAH 4.1.0 and linked against
SPheno 3.2.4).  FlexibleSUSY and Softsusy were compiled with g++ 4.8.0
and gfortran 4.8.0.  SPheno and SPhenoMSSM were compiled with Intel
ifort 13.1.3 20130607\footnote{We are using Intel's ifort compiler,
  instead of gfortran 4.8.0, because it decreases the run-time of
  SPheno and SPhenoMSSM by approximately a factor $1.5$.}.  We're
generating random CMSSM parameter points with $m_0\in [50,1000]$
$m_{1/2}\in [50,1000]$, $\tan\beta\in [1,100]$, $\sign\mu\in
\{-1,+1\}$ and $A_0\in [-1000,1000]$.  The following SLHA template
file was used:
\begin{verbatim}
Block MODSEL                 # Select model
    6    0                   # flavour violation
    1    1                   # mSUGRA
Block SMINPUTS               # Standard Model inputs
    1   1.279180000e+02      # alpha^(-1) SM MSbar(MZ)
    2   1.166390000e-05      # G_Fermi
    3   1.189000000e-01      # alpha_s(MZ) SM MSbar
    4   9.118760000e+01      # MZ(pole)
    5   4.200000000e+00      # mb(mb) SM MSbar
    6   1.709000000e+02      # mtop(pole)
    7   1.777000000e+00      # mtau(pole)
Block SOFTSUSY               # SOFTSUSY specific inputs
    1   1.000000000e-04      # tolerance
    2   2                    # up-quark mixing (=1) or down (=2)
    3   0                    # printout
    5   1                    # 2-loop running
    7   2                    # EWSB and Higgs mass loop order
Block FlexibleSUSY
    0   1.000000000e-04      # precision goal
    1   0                    # max. iterations (0 = automatic)
    2   0                    # algorithm (0 = two_scale, 1 = lattice)
    3   0                    # calculate SM pole masses
    4   1                    # calculate loop masses
    5   1                    # EWSB loop order
Block SPhenoInput            # SPheno specific input
    1  -1                    # error level
    2   1                    # SPA conventions
    11  0                    # calculate branching ratios
    13  0                    # include 3-Body decays
    12  1.000E-04            # write only branching ratios larger than this value
    31  -1                   # fixed GUT scale (-1: dynamical GUT scale)
    32  0                    # Strict unification
    34  1.000E-04            # Precision of mass calculation
    35  40                   # Maximal number of iterations
    37  1                    # Set Yukawa scheme
    38  2                    # 1- or 2-Loop RGEs
    50  1                    # Majorana phases: use only positive masses
    51  0                    # Write Output in CKM basis
    52  0                    # Write spectrum in case of tachyonic states
    55  1                    # Calculate one loop masses
    57  0                    # Calculate low energy constraints
    60  0                    # Include possible, kinetic mixing
    65  1                    # Solution tadpole equation
    75  0                    # Write WHIZARD files
    76  0                    # Write HiggsBounds file
    86  0.                   # Maximal width to be counted as invisible in Higgs decays
    510 0.                   # Write tree level values for tadpole solutions
    515 0                    # Write parameter values at GUT scale
    520 0.                   # Write effective Higgs couplings (HiggsBounds blocks)
    525 0.                   # Write loop contributions to diphoton decay of Higgs
Block MINPAR
    1   [50..1000]           # m0(MX)
    2   [50..1000]           # m12(MX)
    3   [1..100]             # tan(beta)(MZ) DRbar
    4   {-1,+1}              # sign(mu)
    5   [-1000..1000]        # A0(MX)
\end{verbatim}
The SLHA file was passed to each spectrum generator and the
(wall-clock) time was measured until the program has finished.  The
average run-times for three different CPU types can be found in
\tablename~\ref{tab:run-time-comparison}.  The first row shows the
run-time on a Intel Core2 Duo (P8600) where only one core was enabled.
The second row shows the run-time on a Intel Core2 Duo (P8600) where
both cores were enabled.  In the third row a Intel Xeon (L5640) was
used, which has $6$ CPU cores.
\begin{table}[tbh]
  \centering
  \begin{tabular}{llll}
    \toprule
                          & Intel Core2 Duo            & Intel Core2 Duo             & Intel Xeon\\
                          & (1 core, $2.40\unit{GHz}$) & (2 cores, $2.40\unit{GHz}$) & (6 cores, $2.27\unit{GHz}$)\\
    \midrule
    FlexibleSUSY 0.5.3    & $0.151\unit{s}$    & $0.114\unit{s}$   & $0.081\unit{s}$\\
    Softsusy 3.4.0        & $0.178\unit{s}$    & $0.174\unit{s}$   & $0.157\unit{s}$\\
    SPheno 3.2.4          & $0.120\unit{s}$    & $0.118\unit{s}$   & $0.108\unit{s}$\\
    SPhenoMSSM 4.1.0${}^*$ & $0.463\unit{s}$    & $0.450\unit{s}$   & $0.403\unit{s}$\\
    \bottomrule
  \end{tabular}
  \caption{Average run-time for random CMSSM parameter points.
    FlexibleSUSY and Softsusy were compiled
    with g++ 4.8.0 and gfortran 4.8.0.  SPheno and SPhenoMSSM were
    compiled with ifort 13.1.3 20130607.
    ${}^*$ SPhenoMSSM was generated with SARAH 4.1.0 and linked
    against SPheno 3.2.4.}
  \label{tab:run-time-comparison}
\end{table}
We find that on the single-core machine SPheno is fastest with
$0.120\unit{s}$, and FlexibleSUSY is second with $0.151\unit{s}$.  In
the case of 2 or 6 CPU cores, FlexibleSUSY is fastest with
$0.114\unit{s}$ and $0.081\unit{s}$, respectively.  The reason is that
FlexibleSUSY calculates each pole mass in a separate thread, and
therefore benefits from multi-core CPUs.

\appendix
\section{Examples}
\subsection{NMSSM}
Detailed steps to create NMSSM model file.
\begin{enumerate}
\item Create new directory \code{mkdir myNMSSM}
\item Move into this new directory \code{cd myNMSSM}
\item copy the MSSM model file in \code{cp /path/to/SARAH/Models/MSSM/MSSM.m  myNMSSM.m}
\item Edit myNMSSM.m as follows
\begin{enumerate}
\item  Change Model name, author and date
\item Add singlet superfield to SuperFields
\begin{lstlisting}
SuperFields[[8]] = {s, 1, sR,    0, 1,  1, RpM};
\end{lstlisting}
\item Modify superpotential
\begin{lstlisting}
SuperPotential = Yu u.q.Hu - Yd d.q.Hd - Ye e.l.Hd + \[Lambda] s.Hu.Hd +  \[Kappa]/3 s.s.s ;
\end{lstlisting}
\item Add new VEV
\begin{lstlisting} 
  DEFINITION[EWSB][VEVs]= 
  { {SHd0, {vd, 1/Sqrt[2]}, {sigmad, \[ImaginaryI]/Sqrt[2]},{phid,1/Sqrt[2]}},
    {SHu0, {vu, 1/Sqrt[2]}, {sigmau, \[ImaginaryI]/Sqrt[2]},{phiu,1/Sqrt[2]}},
    {SsR,  {vs, 1/Sqrt[2]}, {sigmaS, \[ImaginaryI]/Sqrt[2]},{phiS,1/Sqrt[2]}}
};
\end{lstlisting}
\item  Add the singlino to the list of gauge eigenstate dirac fermions
\begin{lstlisting}
  DEFINITION[GaugeES][DiracSpinors]={
 .............
  FS -> {FsR,conj[FsR]}   
};
\end{lstlisting}
\item Edit mass eigenstates to give how the gauge eigenstates mix into mass eigenstates.  Here we add to the cp even higgs a real singlet, to the CP odd the imaginary part of the singlet and finally to the neutralinos we add the singlino.
